Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data

### 2021-01-19 documents/ovando-etal-assessing-global-fisheries.Rmd
back to master

 tmp %>% 
  knitr::kable(caption = "Estimates of B/B~MSY~ by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database. All analysis were conducted in the R programming language [@rcoreteam2019]. Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries). We encourage readers to explore the documentation available at the package website at www.github.com/danovando/sraplus. Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo []. Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results that are not based on an index of abundance are estimated through stock reduction analysis. All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.  For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

@@

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus. We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database. We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries. For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus (See Supplementary Information). All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.

@@

We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available. We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values. While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).

@@

```{r cs-plot, fig.cap="RAM values of B/B~MSY~ and F/F~MSY~ (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%. Each point is a stock. Black dashed line shows the 1:1 relationship."}
ex_scatter_plot
```

@@

We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status. Overall the sraplus estimates of B/B~MSY~ resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves. This becomes clearer once we consider the performance of sraplus models fit to combinations of broadly available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The status of most stocks in RAM is also relatively good, with recent B/B~MSY~ values generally near one. This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.

@@

```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent B/B~MSY~ by FAO statistical area from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
ram_mpe_map_plot
```

```{r mape-map, fig.cap = "Median absolute percent error in most recent B/B~MSY~ by FAO statistical area from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}
ram_mape_map_plot
```

```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent B/B~MSY~ of 0.4,1, or 1.6."}

@@

```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of B/B~MSY~ in the most recent 5 years of data available for each model fit. Distribution is full posterior probability distribution. Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th. Change is relative to the mean performance of a catch-only heuristic model."}
b_voi_plot + 
  geom_vline(aes(xintercept = 0), alpha = 0.75)
```

>>>>

 tmp %>% 
  knitr::kable(caption = "Estimates of *B/B~MSY~* by FAO, RAM Legacy Stock Assessment Database through Hilborn et al. 2020, Costello et al. 2016, 
               Rosenberg et al. 2018, and Pauly 2007",
               booktabs = TRUE,
               format = "pipe") %>% 
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)

@@

We compared estimates of stock status, defined by *B/B~MSY~* obtained from sraplus using various forms of broadly available data to the best-available estimates of *B/B~MSY~* available for those fish populations reported in the RAM database. All analysis were conducted in the R programming language [@rcoreteam2019]. Model fitting was conducted using Rcpp [@eddelbuettel2011] and Stan [@carpenter2017] implemented through Template Model Builder [@kristensen2016] by the tmbstan package [@monnahan2018]. The sraplus package is publicly available at [github.com/danovando/sraplus](https://github.com/DanOvando/sraplus), and all materials needed to fully reproduce this manuscript are available at [github.com/DanOvando/assessing-global-fisheries](https://github.com/DanOvando/assessing-global-fisheries). Here we describe the structure of the population model underpinning sraplus, the estimation models used, and the construction of priors used in this paper.

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo with the No-U-Turn sampler [@hoffman2011]. Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, *B/B~MSY~*, *F*, or *F/F~MSY~*. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results that are not based on an index of abundance are estimated through stock reduction analysis. All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.  For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

@@

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus. We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database. We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries. For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on *B/B~MSY~* or *F/F~MSY~* in the most recent year of the fishery usable by sraplus (See Supplementary Information). All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.

@@

We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available. We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values. While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both *B/B~MSY~* and *F/F~MSY~* (Fig.\@ref(fig:cs-plot)).

@@

```{r cs-plot, fig.cap="RAM values of *B/B~MSY~* and *F/F~MSY~* (x-axes) for case study fisheries plotted against estimated values (y-axes) using CMSY [@froese2017], priors informed by stock-specific Fisheries Management Index (FMI) and swept area ratio (SAR) scores, and an abundance index based on reconstructed effort trends assuming a rate of technological increase of 2.6%. Each point is a stock. Black dashed line shows the 1:1 relationship."}
ex_scatter_plot
```

@@

We next assessed the ability of FMI, SAR, and effort data to improve estimates of global stock status. Overall the sraplus estimates of *B/B~MSY~* resulting from using the RAM data were relatively accurate and unbiased at a macro level (MPE `r percent(perf$mpe[perf$"Data Used" == "ram-data"])`,MAPE `r percent(perf$mape[perf$"Data Used" == "ram-data"])`, accuracy = `r percent(perf$accuracy[perf$"Data Used" == "ram-data"])`, Table.\@ref(tab:perf-tab), Fig.\@ref(fig:mpe-map)-\@ref(fig:acc-map)). This exercise tells us that given sufficiently high quality data, a surplus production model such as sraplus is reasonably capable of reproducing the global state of fisheries as understood from formally assessed fisheries.

Performance limitations in assessments of global fishery status then are likely to arise less from model misspecification than from the quality of the data themselves. This becomes clearer once we consider the performance of sraplus models fit to combinations of broadly available datasets. Many of the datasets used produced similar levels of bias as the RAM data (Table.\@ref(tab:perf-tab)). However, this is somewhat an artifact of the data. The status of most stocks in RAM is also relatively good, with recent *B/B~MSY~* values generally near one. This means that a model that more or less reproduces the global average of stock status will be relatively unbiased on average, but imprecise. Focusing on MAPE instead, the error of the models jumps dramatically as soon as data other than RAM are used, to a minimum value of `r percent(perf$mape[2])` and a maximum of `r percent(max(perf$mape))`. The mean accuracy of the sraplus models across all non-RAM data fits was only `r percent(mean(perf$accuracy[!perf$"Data Used" %in% c("ram-data","guess")]))`. Note that there are only three bins in the FAO stock status classifications, and a "model" that randomly assigns a stock to a status category would have a mean accuracy of `r percent(perf$accuracy[perf$"Data Used" == "guess"])`.

@@

```{r mpe-map, fig.width=8, fig.height=8, fig.cap="Median percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
ram_mpe_map_plot
```

```{r mape-map, fig.cap = "Median absolute percent error in most recent *B/B~MSY~* by FAO statistical area from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. RAM U/U~MSY~ assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}
ram_mape_map_plot
```

```{r acc-map, fig.cap="Mean classification accuracy (assignment to FAO stock status category) by FAO statistical area arising from different data sources. RAM Index refers to catch and abundance index drawn from RAM. Effective CPUE refers to an index of abundance based on reconstructed effort data. Effective CPUE+ uses CPUE along with Fisheries Management Index (FMI) and/or swept area ratio (SAR) data. For both CPUE series 'nominal' assumes a 0% technology creep, for effective a 2.6% technology creep is assumed. RAM *U/U~MSY~* assumes all fisheries in the region share a common U/U~MSY~ series with formally assessed fisheries in the region. FMI uses FMI scores to develop a prior on recent fishing mortality rates, SAR does the same but based on swept area ratio. CMSY uses the methods from Froese et al. 2017 [@froese2017]. Guess assignes a random recent *B/B~MSY~* of 0.4,1, or 1.6."}

@@

```{r voi-plot, fig.cap="Posterior probability distributions of estimated effect of different data types on root mean squared error of *B/B~MSY~* in the most recent 5 years of data available for each model fit. Distribution is full posterior probability distribution. Point is median, thicker black section inner 66th quantile of the posterior, the thinner black line the 95th. Change is relative to the mean performance of a catch-only heuristic model."}
b_voi_plot + 
  geom_vline(aes(xintercept = 0), alpha = 0.75)
```
